CorReg: model-based pretreatment for regression with correlated variables. Application in steel industry

### 2014-04-01 www/article13.tex
up MQE
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@487 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\subsection{Estimator and properties}
	The marginal regression model simply is
	\begin{equation}
		Y_{|X_1}= X_1\alpha_1+ \varepsilon_{\alpha}
	\end{equation}			
		So we have the \textsc{OLS} unbiased estimator of $\alpha$: 
		\begin{equation}
			\hat{\alpha}_{1} = (X'_{1} X_{1})^{-1}X'_{1}Y  \textrm{ and }\hat{\alpha}_{2} = 0
		\end{equation}
		We see in (\ref{Trueexpl}) that it gives an unbiased estimation of $Y$ and $\alpha$
		but in terms of $\beta$ this estimator is biased:
		\begin{equation}
			\operatorname{E}[\hat{\alpha_1}|X_1]=\beta_1+\sum_{j \in I_2}\beta_{j}\delta_j \textrm{ and }\operatorname{E}[\hat{\alpha_2}|X_1]=0
		\end{equation}
		with variance:
		\begin{equation}
			\operatorname{Var}[\hat{\alpha}_{1}|X_1]= (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 )(X'_{1} X_{1})^{-1}  \textrm{ and }\operatorname{Var}[\hat{\alpha}_{2}|X_1]= 0 
		\end{equation}
		We see that the variance is reduced compared to OLS described in equation (\ref{eqOLS})(no correlations and smaller matrix give better conditioning ) for small values of $\sigma_j$ $i.e.$ strong correlations.					
			%There is no theoretical guarantee that our model is better. It's just a compromise between numerical issues caused by correlations for estimation and selection versus increased variability due to structural hypothesis. Therefore we made some simulations to compare both methods (see the end of this paper).
			 \\

	To better illustrate the bias-variance compromise, we look at a simple example with $p=3$ variables. $X_1$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $X_3=x_1+x_2+\varepsilon_3$ where $\varepsilon_3\sim{\mathcal{N}(0,\sigma_3^2)}$. We also have $\beta=(1,1,1)$ and $\sigma_Y=10$. Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg} model: $MSE(\hat{\beta}|\beta,X)=Bias^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))$ for several values of $\sigma_3$ (strength of the sub-regression) and $n$. Figures \ref{MQE1} to \ref{MQE3} show the theoretical MSE evolution with the strength of the sub-regression:
	\begin{equation}
		1-\mathcal{R}^2=\frac{\operatorname{Var}(\varepsilon_3)}{\operatorname{Var}(x_3)}=\frac{\sigma_3^2}{\sigma_3^2+2}
	\end{equation}

\begin{figure}[h!]
	\begin{minipage}[l]{.32\linewidth}
			\includegraphics[width=170px]{figures/MQEn15sigmaY10.png} 
			\caption{For $n=15$. Dotted: \textsc{Correg}, plain: OLS}\label{MQE1}
	\end{minipage} \hfill
	\begin{minipage}[c]{.32\linewidth}
			\includegraphics[ width=170px]{figures/MQEn100sigmaY10.png} 
			\caption{For $n=100$. Dotted: \textsc{Correg}, plain: OLS}
	\end{minipage} \hfill
   \begin{minipage}[r]{.32\linewidth}
			\includegraphics[width=170px]{figures/MQEn1000sigmaY10.png} 
			\caption{For $n=1000$. Dotted: \textsc{Correg}, plain: OLS.} \label{MQE3}
   \end{minipage} 
\end{figure} 
It is clear in Figures \ref{MQE1} to \ref{MQE3} that the reduced model is more robust than OLS. And even when sub-regression are week (small value of $1-\mathcal{R}^2$) it does not explode. We also see that the error implied by strong correlations shrinks whit the rise of $n$.  $MSE_{OLS}=0+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))=\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{1}))+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{2}))$. both terms are multiplied by $\sigma_Y^2$ and so is $\operatorname{Tr}(\operatorname{Var}(\hat{\alpha}_{1}))$. Thus, as $\operatorname{Var}(\hat{\alpha}_{1})=0$ when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus OLS. It illustrates the importance of dimension reduction when the model has a strong noise (very usual case on real datasets where true model is not even exactly linear).

@@

\section{Numerical results on simulated datasets} \label{sectionsimul}
	\subsection{The datasets}	
	Now we have defined the model and the way to obtain it, we can have a look on some numerical results to see if \textsc{CorReg} 	keeps its promises.
	The \textsc{CorReg} package has been tested on simulated datasets. 
Section \ref{compZ} shows the results obtained in terms of $\hat{S}$. Sections \ref{tableMSEsimtout} and \ref{tableMSEsimgauche} show the results obtained using only \textsc{CorReg}, or \textsc{CorReg} combined with other methods. Tables give both mean and standard deviation of the observed Mean Squared Errors (MSE) on a validation sample of $1 000$ individuals. For each simulation,  $p=40$, $\sigma_Y=10$, $\sigma=0.001$, variables $X_1$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda$ and which standard deviation also is $\lambda$. The $\delta_j$ and $\beta_j$ are generated according to the same Poisson law but with a random sign. $S$ only contains binary relationships but \textsc{CorReg} was only constrained to $\max (\hat{p}_1^j)=5$.  
	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the walk was launched on $20$ initial structures with a maximum of 9 000 steps each time.

>>>>

\subsection{Estimator and properties}
	The marginal regression model simply is
	\begin{equation}
		Y_{|X_1}= X_1\alpha_1+ \varepsilon_{\alpha}
	\end{equation}			
		So we have the \textsc{OLS} unbiased estimator of $\alpha$: 
		\begin{equation}
			\hat{\alpha}_{1} = (X'_{1} X_{1})^{-1}X'_{1}Y  \textrm{ and }\hat{\alpha}_{2} = 0
		\end{equation}
		We see in (\ref{Trueexpl}) that it gives an unbiased estimation of $Y$ and $\alpha$
		but in terms of $\beta$ this estimator is biased:
		\begin{equation}
			\operatorname{E}[\hat{\alpha_1}|X_1]=\beta_1+\sum_{j \in I_2}\beta_{j}\delta_j \textrm{ and }\operatorname{E}[\hat{\alpha_2}|X_1]=0
		\end{equation}
		with variance:
		\begin{equation}
			\operatorname{Var}[\hat{\alpha}_{1}|X_1]= (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 )(X'_{1} X_{1})^{-1}  \textrm{ and }\operatorname{Var}[\hat{\alpha}_{2}|X_1]= 0 
		\end{equation}
		We see that the variance is reduced compared to OLS described in equation (\ref{eqOLS})(no correlations and smaller matrix give better conditioning ) for small values of $\sigma_j$ $i.e.$ strong correlations. So we play on the bias-variance tradeoff, reducing the variance by adding a bias.		
			%There is no theoretical guarantee that our model is better. It's just a compromise between numerical issues caused by correlations for estimation and selection versus increased variability due to structural hypothesis. Therefore we made some simulations to compare both methods (see the end of this paper).
			 \\

@@

\section{Numerical results on simulated datasets} \label{sectionsimul}

	\subsection{Theoretical Mean squared errors on $\hat \beta$} 
	To better illustrate the bias-variance tradeoff, we look at a simple example with $p=3$ variables. $X_1$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $X_3=x_1+x_2+\varepsilon_3$ where $\varepsilon_3\sim{\mathcal{N}(0,\sigma_3^2)}$. We also have $\beta=(1,1,1)$ and $\sigma_Y=10$. Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg} model: $MSE(\hat{\beta}|\beta,X)=Bias^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))$ for several values of $\sigma_3$ (strength of the sub-regression) and $n$. Figures \ref{MQE1} to \ref{MQE3} show the theoretical MSE evolution with the strength of the sub-regression:
	\begin{equation}
		1-\mathcal{R}^2=\frac{\operatorname{Var}(\varepsilon_3)}{\operatorname{Var}(x_3)}=\frac{\sigma_3^2}{\sigma_3^2+2}
	\end{equation}

\begin{figure}[h!]
	\begin{minipage}[l]{.32\linewidth}
			\includegraphics[width=170px]{figures/MQEn15sigmaY10.png} 
			\caption{For $n=15$. Dotted: \textsc{Correg}, plain: OLS}\label{MQE1}
	\end{minipage} \hfill
	\begin{minipage}[c]{.32\linewidth}
			\includegraphics[ width=170px]{figures/MQEn100sigmaY10.png} 
			\caption{For $n=100$. Dotted: \textsc{Correg}, plain: OLS}
	\end{minipage} \hfill
   \begin{minipage}[r]{.32\linewidth}
			\includegraphics[width=170px]{figures/MQEn1000sigmaY10.png} 
			\caption{For $n=1000$. Dotted: \textsc{Correg}, plain: OLS.} \label{MQE3}
   \end{minipage} 
\end{figure} 
It is clear in Figures \ref{MQE1} to \ref{MQE3} that the reduced model is more robust than OLS. And even when sub-regression are week (small value of $1-\mathcal{R}^2$) it does not explode. We also see that the error implied by strong correlations shrinks whit the rise of $n$.  $MSE_{OLS}=0+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))=\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{1}))+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{2}))$. both terms are multiplied by $\sigma_Y^2$ and so is $\operatorname{Tr}(\operatorname{Var}(\hat{\alpha}_{1}))$. Thus, as $\operatorname{Var}(\hat{\alpha}_{2})=0$ when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus \textsc{OLS}. It illustrates the importance of dimension reduction when the model has a strong noise (very usual case on real datasets where true model is not even exactly linear).

	\subsection{The datasets}	
	Now we have defined the model and the way to obtain it, we can have a look on some numerical results to see if \textsc{CorReg} 	keeps its promises.
	The \textsc{CorReg} package has been tested on simulated datasets. 
Section \ref{compZ} shows the results obtained in terms of $\hat{S}$. Sections \ref{tableMSEsimtout} and \ref{tableMSEsimgauche} show the results obtained using only \textsc{CorReg}, or \textsc{CorReg} combined with other methods. Tables give both mean and standard deviation of the observed Mean Squared Errors (MSE) on a validation sample of $1 000$ individuals. For each simulation,  $p=40$, $\sigma_Y=10$, $\sigma=0.001$, variables $X_1$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda$ and which standard deviation also is $\lambda$. The $\delta_j$ and $\beta_j$ are generated according to the same Poisson law but with a random sign. $S$ only contains binary relationships but \textsc{CorReg} was only constrained to $\max (\hat{p}_1^j)=5$.  
	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the walk was launched on $20$ initial structures with a maximum of 9 000 steps each time.
